A tool for designing digital filters for the Hankel and Fourier transforms in potential, diffusive, and wavefield modeling

### 2018-11-21 ManuscriptGeophysics2018.tex
Move // inside \old{}
Within preparation for final manuscript

%%fakesection ===    PACKAGES & DEFINITIONS    ===
% \usepackage{showframe}

%% ~ ADDITIONAL PACKAGES TO GEOPHYSICS.CLS
\usepackage[USenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[strings]{underscore}
\usepackage[pdftex, final]{hyperref}
\hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}

@@

% Own commands
\usepackage{xspace}
\newcommand{\empymod}{\texttt{empymod}\xspace}
\newcommand{\mr}[1]{\mathrm{#1}}

@@

\address{
\footnotemark[1]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft,
E-mail: \href{mailto:dieter@werthmuller.org}{Dieter@Werthmuller.org};
\href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
\footnotemark[2]Instituto Mexicano del Petróleo,
Eje Central Lázaro Cárdenas Norte 152,
Col. San Bartolo Atepehuacan C.P. 07730,
Ciudad de México, México,
\footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
305C Oceanography,
61 Route 9W, PO Box 1000,
Palisades NY 10964-8000 US,
E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
}

@@

\cite{PhD.70.Ghosh} derived the filter coefficients in the spectral domain by
dividing the output spectrum by the input spectrum followed by an inverse
Fourier transform. Improvements to the determination of filter coefficients
were provided by \cite{EXG.75.ONeill, GEO.77.Nyman, GEO.82.Das}, or
specifically for the Fourier transform by \cite{GP.86.Nissen}. A direct
integration method was used by \cite{GP.76.Bichara} and \cite{GP.78.Bernabini}.
\cite{GP.79.Koefoed} proposed a Wiener-Hopf least-squares method, which was
further improved by many authors \citep{GP.82.Guptasarma, GEO.1982.Murakami,
GP.97.Guptasarma}. \cite{GP.07.Kong} proposed a direct matrix inversion method
to solve the convolution equation, which requires only the input and output
sample values. To evaluate the filter's effectiveness, he defines a good filter
as one that recovers small or weak diffusive EM fields. This method was also
used by \cite{GEO.09.Key, GEO.12.Key} to create a suite of filters accurate for
marine EM data. Most works have published filters for the Hankel transform with
$J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if applied to the
Fourier sine/cosine transform), as all higher Bessel functions can be
rewritten, via recurrence relations, using only use these two.
\cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
weights.

@@

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
was even thought that the filter method works only in the diffusive limit where
the quasistatic approximation is valid \citep[e.g.,][]{GEO.15.Hunziker}.
Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
center-frequency GPR data using a 401 point filter that was designed for
diffusive EM, and found that it was orders of magnitude faster than quadrature
approaches. However, good quality results were only obtained for the first
arrivals at short time-offsets and the later arrivals had poor quality. Yet,
this promising initial test \old{suggested}\new{suggests} it might be possible
to specifically design a special filter for GPR frequencies where the EM fields
propagate as waves rather than by diffusion.

@@

In the following sections we give a brief overview of the methodology, the
theory, and the code, and then show examples of its usage, as well as
presenting new filters for both CSEM and GPR data. The current article falls
therefore into all three mentioned categories by the new application of DLF to
wave phenomena, \new{which means that the DLF can probably be used to model
seismic wavefields as well,} and by providing new filters with the accompanying
code. The algorithm and many more examples of its usage can be found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

@@

The main features of \texttt{fdesign} and some differences between it and the
methods it is based upon are:
\begin{itemize}
  \item The algorithm computes many different filters for various spacing and
    shift values (brute-force), with an optional minimization routine for
    the best outcome of the brute-force step.
  \item \cite{GP.07.Kong} and \cite{GEO.12.Key} optimize a filter for the $J_0$
    filter, and then use the obtained shift and spacing value to calculate the
    $J_1$ values. The presented algorithm can optimize either $J_0$, $J_1$, or
    both in the optimization process. Five transform-pairs for each $J_0$ and
    $J_1$, and three transform pairs for each sine and cosine of often used
    functions are included in the routine, but any other transform pair can be
    provided as input.
  \item More complex or more specific numerical transform  pairs  can be
    provided instead of theoretical transform pairs by using the modeler
    \texttt{empymod}.
  \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
    \emph{weak} diffusive EM fields. In  \texttt{fdesign}, you can define a
    relative error level that defines up to what error an obtained result is
    good or not. The optimization seeks to find the smallest amplitude field
    that is accurately modeled (i.e., minimizing the amplitude), but it has
    additionally another mode, where it will maximize the output abscissae $r$
    of the transform pair instead of minimizing the amplitude. The two modes
    yield the same result in the general, simple case of fast decaying
    transform pairs. However, for some complex models and specifically for very
    high frequencies, maximizing $r$ will yield much better and more consistent
    results.
  \item The algorithm can solve \old{under-,} equal-\old{,} and over-determined
    linear systems.
  \item The real or the imaginary part can be used for the inversion of complex
    transform pairs.
\end{itemize}

@@

In the direct matrix inversion method for solving for the digital filter
coefficients, equation~\ref{eq:filtapprox} is cast as a linear system\old{ as
follows.}\new{, where} $\Delta$ and $\delta$ are preassigned scalar values and
$r$ is a range of $M$ preassigned values $r_m$. The filter base coefficients
$b_n$ are computed using equation~\ref{eq:base} and an array of values for $l$
are computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
matrix $\bf{A}$ has dimensions $M \times N$ with coefficients \old{$A_{mn} =
f(l_{mn})/r_n$}\new{$A_{mn} = f(l_{mn})/r_m$} and the right hand side (RHS)
vector $\bf{v}$ has elements \old{$v_n = F(r_n)$}\new{$v_m = F(r_m)$}. The
resulting linear system
%
\begin{equation}
  \bf{A}\bf{h} = \bf{v}
  \label{eq:,linearSystem}
\end{equation}
%
can be solved by direct matrix inversion\new{, or any other matrix inversion
routine,} to obtain the vector of filter coefficients $\bf{h}$. For a given
filter length $N$, there are many subjective choices that go into the practical
implementation of this method, including the choice of the transform pairs
$F(r)$ and $f(l)$, as well as the value for $M$ and the particular range and
spacing of the values $r_m$. Once values are chosen for these variables, an
optimal filter can be found by a grid search over $\Delta$ and $\delta$ for
values that produce a high quality filter. The choice of metric for what
constitutes a high quality filter is also subjective and is further discussed
below.

@@

The basic steps are as follows:
\begin{enumerate}
  \item Evaluate RHS of check-function $f_\mr{C}$:\newline
    \old{$d_\mr{R} = f_\mr{C}.\mr{rhs}(r)$}\\
    \new{$d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$}
  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force\new{;
      parameters on the left-hand side are overwritten where $i$, $j$
      are not mentioned}):
    \begin{enumerate}
      \item Calculate filter base:\newline
        \old{$b_n = \exp\left\{\Delta_i(-N/2+n) + \delta_j\right\}$}\\
        \new{$b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
      \delta_j\right\}$}
      \item Get required RHS and LHS evaluation points:\newline
        $r_\mr{I} = p(b, N)$\newline
        $l_\mr{I} = b/r_\mr{I}$
      \item Invert for filter coefficients:\newline
        \old{$J_{ij} = \mr{solve}(f_\mr{I}.\mr{lhs}(l_\mr{I}),
        f_\mr{I}.\mr{rhs}(r_\mr{I}))$}\\
        \new{$J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$}
    \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
      filter \old{$J_{ij}$}\new{$J$}:\newline
      \old{$d_\mr{F} = f_\mr{C}.\mr{lhs}(b/r)\cdot J_{ij}\ /\ r$}\\
      \new{$d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$}
      \item Store minimum recovered amplitude or maximum $r$ where relative
        error is less than the provided error:\newline
        $\chi_{ij} = g\left[
        \mr{argmin}\left(|(d_\mr{F}-d_\mr{R})/d_\mr{R}| >
        \mr{error}\right) -1\right]\ $, where $g$ is either $d_\mr{R}$
        or $1/r$.
    \end{enumerate}
  \item \old{Return filter}\new{Re-calculate (steps 2a-2c) filter base and}
    coefficients which yield minimum amplitude or maximum $r$ (a local
    minimization can be run to polish the brute-force result):\newline
    \new{$i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$}\newline
    \new{recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$}\newline
    \old{return $J\left[\mr{argmin}(\chi)\right] $}\newline
  \item \new{Return filter base and coefficients and error matrix (for visual
    QC):}\newline
    \new{return $b$, $J$, $\chi$}\newline
\end{enumerate}

\new{To solve step 2(c) we implemented the QR-decomposition method in
  \texttt{fdesign}.}

\subsection{Minimization criteria}
Figure~\ref{fig:InvCrit} shows the differences between the different
minimization approaches; in (a) for a conventional, fast decaying transform
pair \new{(using the $J_0$ kernel of a fullspace with vertical distance between
source and receiver of 50\,m, frequency $f=1\,$Hz, and resistivity
$\rho=1\,\Omega\,$m)}; and in (b) for a more complex, high frequency layered
model \new{(using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
in the section \emph{Ground-penetrating radar}, with receiver located at 1\,m
depth)}. The circles show the minimum amplitude used in previous approaches.
This criterion is not ideal, as it is subject to some random fluctuations and
depends on the choice of $r$. The squares show the minimum amplitude given a
certain acceptable error, and the diamonds show the maximum $r$ given a certain
acceptable error (the inversion is a minimization process, it therefore
minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
acceptable error and the maximum $r$ given an acceptable error will yield the
same result. However, in complex cases the maximum $r$ is more consistent and
therefore a much better criterion.
%
\plot*{InvCrit}{width=.9\textwidth}{(a) A regular RHS curve of a transform pair
  with a purely decaying function. Minimum amplitude or maximum $r$ yield the
  same result in this scenario. Using a relative error criteria is more stable
  than just the absolute minimum amplitude. (b) A 1D model for $f=200\,$MHz.
  Here, the maximum $r$ provides a better criteria for the inversion.}
%

@@

\subsection{Design}
\old{Figure~\ref{fig:201b} shows the solution spaces of four consecutive
inversion runs, where each run is a focus on a subsection of the previous run,
indicated by the red square. The filter length in this case is $N=201$, and the
theoretical transform pairs are as given, e.g., in
}\cite{USGS.75.Anderson}\old{:}
\new{Figures~\ref{fig:201a} and~\ref{fig:201b} show the solution spaces of
seven consecutive inversion runs using the following theoretical transform
pairs} \citep[e.g.,][]{USGS.75.Anderson}\new{,}
%
\begin{align}
  \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
  \frac{\exp\left(\frac{-r^2}{4a}\right)}{2a}\ , \\
  \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
  \frac{r}{4a^2} \exp\left(-\frac{r^2}{4a}\right)\ ,
  \label{eq:j01}
\end{align}
%
where $a$ was set to 5. In the algorithm you can provide one pair
for the inversion, and a different pair to get the minimum amplitude or the
maximum $r$. In this case the same transform pair was used for both the
inversion and the check of quality. The acceptable relative error was set to
1\,\%. The RHS evaluation parameter for the inversion was set so  that an
over-determined system was evaluated with $M = 2N$ equations, where $r_\mr{I}$
was logarithmically spaced from $\log_{10}[1/\max(b)] - 1$ to
$\log_{10}[1/\min(b)] + 1$, where $b$ are the filter abscissae as given in
equation~\ref{eq:base}. This is accomplished in \texttt{fdesign} by setting
parameter $r_\mr{def} = (1, 1, 2)$.

\new{Figure~\ref{fig:201a} shows three different inversion runs for $N=101$,
201, and 401 for a wide range of spacing and shift values.
From this one has to decide for a filter length. Longer filters tend to give
more precise results, but their transform is computationally more expensive as
more values have to be evaluated. The length of the filter does not has to be
an odd number, but many published filters do have an odd number of
coefficients. The integral runs from $-\infty$ to $+\infty$ and hence an odd
point filter comes naturally by truncating the integral to a finite sum from
$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value
$\delta = 0$. If it differs from zero then there is no advantage in having an
odd point filter.}
%
\plot*{201a}{width=\textwidth}{%
  \new{Test of different filter lengths. Filter length is a trade-off between
    precision and speed, as shorter filters are faster but less precise, and
  longer filters are more precise but computationally more expensive. The red
  square in subplot (b) is the starting point of Figure~\ref{fig:201b}.} }%
%

\new{Figure~\ref{fig:201b} shows four consecutive inversion runs for $N=201$,
where each run is a focus on a subsection of the previous run, indicated by the
red square.} From the low resolution overview runs (a) and (b) it looks like
this is a standard, straight-forward minimization problem. However, from the
more detailed results in (c) and (d) it becomes obvious that it is a
minimization problem that has to be solved stochastically, as there are
solutions with 2 orders of magnitude difference apparently randomly next to
each other.
%
\plot*{201b}{width=\textwidth}{%
  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
  consecutive run zooms into a portion of the previous solution space,
  indicated by the red square. The more in detail we obtain the solution space,
  the more random appears to be the distribution.}%
%

Figure~\ref{fig:201c} shows in (a) the filter values for $J_0$ and $J_1$ of the
best 201 point filter obtained this way, and in (b) its right-hand-side
solution. The black dots indicate negative values, from which it can seen that
adjacent values are often alternating between positive and negative
contributions.
%
\plot*{201c}{width=.9\textwidth}{(a) Filter values of the best obtained 201\,pt
  filter with the corresponding check of quality in (b). Black points indicate
  negative values, which shows that adjacent values have often opposite signs.}
%

Designing filters is to a large extent trial and error. All input variables
influence the outcome, and often \old{a good}\new{an even better} filter is
found by sheer luck of choosing the right starting parameters\new{, as can be
seen in Figure~\ref{fig:201c} (h). However, once you are in this zone every
filter is a very good filter}. Each of the input parameters has an effect on
the outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$;
functions that decay rapidly are generally better, as noted by earlier authors
\citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
obviously heavily on the spacing and shift values. Another important point is
how you define the right-hand evaluation points of the inversion
($r_\mr{def}$). Evaluating corresponding transform pairs separately or jointly
also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$,
or equally sine, cosine, or sine \& cosine); also if the real or the imaginary
part is used when complex transform pairs or \texttt{empymod} is used; and if
it is inverted for the minimum amplitude or for the maximum $r$.

\subsection{CSEM}
In this section we compare the 201\,pt filter derived in the previous section
to marine CSEM models used in \cite{GP.07.Kong}, \cite{GEO.12.Key}, as well as
for a land EM case.

Figure~\ref{fig:kongf5} compares the derived 201\,pt filter with the two
half-spaces case used by \cite{GP.07.Kong} in his Figure~5. The model consists
of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
$1\,\Omega\,$m half-space. The signal of a $1\,$Hz frequency $x$-directed
electric source 50\,m above the interface is measured at an $x$-directed
electric receiver at the interface.
%
\plot*{kongf5}{width=\textwidth}{(a) Results of different filters and QWE for
  the model of Figure~5 from \cite{GP.07.Kong} with the relative errors shown
  in (b), using the QWE result. The relative error is meaningless from about
  15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
%
In (a) it can be seen that the new 201\,pt filter is able to recover smaller
amplitudes than the 241\,pt filter from \cite{GP.07.Kong}, the 201\,pt filter
from \cite{GEO.12.Key}, and the 801\,pt filter from \cite{TMS.82.Anderson}
(Wer201, Kong241, Key201, and And801, respectively). It behaves equally well as
the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
with relative and absolute tolerance of $10^{-12}$ and $10^{-30}$,
respectively. The relative error is shown in (b), where the QWE result was
taken as \emph{truth}. For offsets greater than roughly 15.5\,km the relative
error becomes meaningless, as the QWE fails itself; this part is greyed out in
the figure.

>>>>

% Additional packages to geophysics.cls
\usepackage[USenglish]{babel}
\usepackage[utf8]{inputenc}
\usepackage{lmodern}
\usepackage[T1]{fontenc}
\usepackage{amssymb, amsmath, amsfonts}
\usepackage{upquote}
\usepackage[strings]{underscore}
% \usepackage[pdftex, final]{hyperref}
% \hypersetup{allcolors=blue, allbordercolors={0 0 .5}, colorlinks=true}
\usepackage[pdftex, hidelinks]{hyperref}

@@

% Own commands
\newcommand{\mr}[1]{\mathrm{#1}}

@@

\address{
\footnotemark[1]TU Delft,
Building 23,
Stevinweg 1 / PO-box 5048,
2628 CN Delft,
E-mail: \href{mailto:D.Werthmuller@tudelft.nl}{D.Werthmuller@tudelft.nl};
\href{mailto:E.C.Slob@tudelft.nl}{E.C.Slob@tudelft.nl};
\footnotemark[2]Instituto Mexicano del Petróleo,
Eje Central Lázaro Cárdenas Norte 152,
Col. San Bartolo Atepehuacan C.P. 07730,
Ciudad de México, México,
\footnotemark[3]Lamont-Doherty Earth Observatory, Columbia University,
305C Oceanography,
61 Route 9W, PO Box 1000,
Palisades NY 10964-8000 US,
E-mail: \href{mailto:KKey@ldeo.columbia.edu}{KKey@ldeo.columbia.edu}.
}

@@

\cite{PhD.70.Ghosh} derived the filter coefficients in the spectral domain by
dividing the output spectrum by the input spectrum followed by an inverse
Fourier transform. Improvements to the determination of filter coefficients
were provided by \cite{EXG.75.ONeill, GEO.77.Nyman, GEO.82.Das}, or
specifically for the Fourier transform by \cite{GP.86.Nissen}. A direct
integration method was used by \cite{GP.76.Bichara} and \cite{GP.78.Bernabini}.
\cite{GP.79.Koefoed} proposed a Wiener-Hopf least-squares method, which was
further improved by many authors \citep{GP.82.Guptasarma, GEO.1982.Murakami,
GP.97.Guptasarma}. \cite{GP.07.Kong} proposed a direct matrix inversion method
to solve the convolution equation, which requires only the input and output
sample values. To evaluate the filter's effectiveness, he defines a good filter
as one that recovers small or weak diffusive EM fields. This method was also
used by \cite{GEO.09.Key, GEO.12.Key} to create a suite of filters accurate for
marine EM data. Most works have published filters for the Hankel transform with
$J_0$ and $J_1$ Bessel functions (or $J_{-1/2}$, $J_{1/2}$ if applied to the
Fourier sine/cosine transform), as all higher Bessel functions can be
rewritten, via recurrence relations, using only\old{use} these two.
\cite{GP.94.Mohsen} is one of the rare cases which provides $J_2$ filter
weights.

@@

All mentioned publications have in common that they were derived for direct
current methods (DC) or low frequency methods, such as time-domain shallow EM
methods (TEM), or controlled-source electromagnetics (CSEM), but not for high
frequency methods such as ground-pe\-ne\-tra\-ting radar (GPR). Generally it
was even thought that the filter method works only in the diffusive limit where
the quasistatic approximation is valid \citep[e.g.,][]{GEO.15.Hunziker}.
Nevertheless, \cite{GEO.17.Werthmuller} tested DLF for modeling 250 MHz
center-frequency GPR data using a 401 \old{point}\new{pt} filter that was
designed for diffusive EM, and found that it was orders of magnitude faster
than quadrature approaches. However, good quality results were only obtained
for the first arrivals at short time-offsets and the later arrivals had poor
quality. Yet, this promising initial test suggests it might be possible to
specifically design a special filter for GPR frequencies where the EM fields
propagate as waves rather than by diffusion.

@@

In the following sections we give a brief overview of the methodology, the
theory, and the code, and then show examples of its usage, as well as
presenting new filters for both CSEM and GPR data. The current article falls
therefore into all three mentioned categories by the new application of DLF to
wave phenomena, which means that the DLF can probably be used to model
seismic wavefields as well, and by providing new filters with the accompanying
code. The algorithm and many more examples of its usage can be found on
\href{https://github.com/empymod/article-fdesign}{github.com/empymod/article-fdesign}.
The examples can be used as templates to design new filters.

@@

The main features of \texttt{fdesign} and some differences between it and the
methods it is based upon are:
\begin{itemize}
  \item The algorithm computes many different filters for various spacing and
    shift values (brute-force), with an optional minimization routine for
    the best outcome of the brute-force step.
  \item \cite{GP.07.Kong} and \cite{GEO.12.Key} optimize a filter for the $J_0$
    filter, and then use the obtained shift and spacing value to calculate the
    $J_1$ values. The presented algorithm can optimize either $J_0$, $J_1$, or
    both in the optimization process. Five transform-pairs for each $J_0$ and
    $J_1$, and three transform pairs for each sine and cosine of often used
    functions are included in the routine, but any other transform pair can be
    provided as input.
  \item More complex or more specific numerical transform  pairs  can be
    provided instead of theoretical transform pairs by using the modeler
    \texttt{empymod}.
  \item \cite{GP.07.Kong} defines a \emph{good} filter as one that recovers
    \emph{weak} diffusive EM fields. In  \texttt{fdesign}, you can define a
    relative error level that defines up to what error an obtained result is
    good or not. The optimization seeks to find the smallest amplitude field
    that is accurately modeled (i.e., minimizing the amplitude), but it has
    additionally another mode, where it will maximize the output abscissae $r$
    of the transform pair instead of minimizing the amplitude. The two modes
    yield the same result in the general, simple case of fast decaying
    transform pairs. However, for some complex models and specifically for very
    high frequencies, maximizing $r$ will yield much better and more consistent
    results.
  \item The algorithm can solve equal- and over-determined
    linear systems.
  \item The real or the imaginary part can be used for the inversion of complex
    transform pairs.
\end{itemize}

@@

In the direct matrix inversion method for solving for the digital filter
coefficients, equation~\ref{eq:filtapprox} is cast as a linear system,
where $\Delta$ and $\delta$ are preassigned scalar values and
$r$ is a range of $M$ preassigned values $r_m$. The filter base coefficients
$b_n$ are computed using equation~\ref{eq:base} and an array of values for $l$
are computed using $l_{mn} = b_n/r_m$. The linear system's left hand side (LHS)
matrix $\bf{A}$ has dimensions $M \times N$ with coefficients
$A_{mn} = f(l_{mn})/r_m$ and the right hand side (RHS)
vector $\bf{v}$ has elements $v_m = F(r_m)$. The
resulting linear system
%
\begin{equation}
  \bf{A}\bf{h} = \bf{v}
  \label{eq:,linearSystem}
\end{equation}
%
can be solved by direct matrix inversion, or any other matrix inversion
routine, to obtain the vector of filter coefficients $\bf{h}$. For a given
filter length $N$, there are many subjective choices that go into the practical
implementation of this method, including the choice of the transform pairs
$F(r)$ and $f(l)$, as well as the value for $M$ and the particular range and
spacing of the values $r_m$. Once values are chosen for these variables, an
optimal filter can be found by a grid search over $\Delta$ and $\delta$ for
values that produce a high quality filter. The choice of metric for what
constitutes a high quality filter is also subjective and is further discussed
below.

@@

The basic steps are as follows:
\begin{enumerate}
  \item Evaluate RHS of check-function $f_\mr{C}$:\newline
    $d_\mr{R} = f_\mr{C}^\mr{rhs}(r)$
  \item Loop over each value $\Delta_i$, $\delta_j$ (brute force;
      parameters on the left-hand side are overwritten where $i$, $j$
      are not mentioned):
    \begin{enumerate}
      \item Calculate filter base:\newline
        $b_n = \exp\left\{\Delta_i(-\lfloor{(N+1)/2\rfloor}+n) +
      \delta_j\right\}$
      \item Get required RHS and LHS evaluation points:\newline
        $r_\mr{I} = p(b, N)$\newline
        $l_\mr{I} = b/r_\mr{I}$
      \item Invert for filter coefficients:\newline
        $J = \mr{solve}\left[f_\mr{I}^\mr{lhs}(l_\mr{I}),
        f_\mr{I}^\mr{rhs}(r_\mr{I})\right]$
    \item Calculate numerically RHS of check-function $f_\mr{C}$ with current
      filter $J$:\newline
      $d_\mr{F} = f_\mr{C}^\mr{lhs}(b/r)\cdot J\ /\ r$
      \item Store minimum recovered amplitude or maximum $r$ where relative
        error is less than the provided error:\newline
        $\chi_{ij} = g\left[
        \mr{argmin}\left(|(d_\mr{F}-d_\mr{R})/d_\mr{R}| >
        \mr{error}\right) -1\right]\ $, where $g$ is either $d_\mr{R}$
        or $1/r$.
    \end{enumerate}
  \item Re-calculate (steps 2a-2c) filter base and
    coefficients which yield minimum amplitude or maximum $r$ (a local
    minimization can be run to polish the brute-force result):\newline
    $i_\mr{best}$, $j_\mr{best}$ = $\mr{argmin}(\chi)$\newline
    recalculate $b$, $J$ with $i_\mr{best}$, $j_\mr{best}$\newline
    \newline
  \item Return filter base and coefficients and error matrix (for visual
    QC):\newline
    return $b$, $J$, $\chi$\newline
\end{enumerate}

To solve step 2(c) we implemented the QR-decomposition method in
  \texttt{fdesign}.

\subsection{Minimization criteria}
Figure~\ref{fig:Figure-1-bw} shows the differences between the different
minimization approaches; in (a) for a conventional, fast decaying transform
pair (using the $J_0$ kernel of a fullspace with vertical distance between
source and receiver of 50\,m, frequency $f=1\,$Hz, and resistivity
$\rho=1\,\Omega\,$m); and in (b) for a more complex, high frequency layered
model (using the $J_0$ kernel for $f=200\,$MHz for the GPR model described
in the section \emph{Ground-penetrating radar}, with receiver located at 1\,m
depth). The circles show the minimum amplitude used in previous approaches.
This criterion is not ideal, as it is subject to some random fluctuations and
depends on the choice of $r$. The squares show the minimum amplitude given a
certain acceptable error, and the diamonds show the maximum $r$ given a certain
acceptable error (the inversion is a minimization process, it therefore
minimizes $1/r$, not $r$). In simple cases the minimum amplitude given an
acceptable error and the maximum $r$ given an acceptable error will yield the
same result. However, in complex cases the maximum $r$ is more consistent and
therefore a much better criterion.
%
\plot*{Figure-1-bw}{width=.9\textwidth}{(a) A regular \new{right-hand-side
  (}RHS\new{)} curve of a transform pair with a purely decaying function.
  Minimum amplitude or maximum $r$ yield the same result in this scenario.
  Using a relative error criteria is more stable than just the absolute minimum
  amplitude. (b) A 1D model for $f=200\,$MHz. Here, the maximum $r$ provides a
  better criteria for the inversion.}
%

@@

\subsection{Design}
Figures~\ref{fig:Figure-2} and~\ref{fig:Figure-3} show the solution spaces of
seven consecutive inversion runs using the following theoretical transform
pairs \citep[e.g.,][]{USGS.75.Anderson},
%
\begin{align}
  \int^\infty_0\,l \exp\left(-al^2\right) J_0(rl)\,dl &=
  \frac{1}{2a} \exp\left(\frac{-r^2}{4a}\right)\ , \\
  \int^\infty_0\,l^2 \exp\left(-al^2\right) J_1(rl)\,dl &=
  \frac{r}{4a^2} \exp\left(-\frac{r^2}{4a}\right)\ ,
  \label{eq:j01}
\end{align}
%
where $a$ was set to 5. In the algorithm you can provide one pair
for the inversion, and a different pair to get the minimum amplitude or the
maximum $r$. In this case the same transform pair was used for both the
inversion and the check of quality. The acceptable relative error was set to
1\,\%. The RHS evaluation parameter for the inversion was set so  that an
over-determined system was evaluated with $M = 2N$ equations, where $r_\mr{I}$
was logarithmically spaced from $\log_{10}[1/\max(b)] - 1$ to
$\log_{10}[1/\min(b)] + 1$, where $b$ are the filter abscissae as given in
equation~\ref{eq:base}. This is accomplished in \texttt{fdesign} by setting
parameter $r_\mr{def} = (1, 1, 2)$.

Figure~\ref{fig:Figure-2} shows three different inversion runs for $N=101$,
201, and 401 for a wide range of spacing and shift values.
From this one has to decide for a filter length. Longer filters tend to give
more precise results, but their transform is computationally more expensive as
more values have to be evaluated. The length of the filter does not has to be
an odd number, but many published filters do have an odd number of
coefficients. The integral runs from $-\infty$ to $+\infty$ and hence an odd
point filter comes naturally by truncating the integral to a finite sum from
$-(N-1)/2$ to $(N-1)/2$. However, this is only true if the shift value
$\delta = 0$. If it differs from zero then there is no advantage in having an
odd point filter.
%
\plot*{Figure-2}{width=\textwidth}{%
  Test of different filter lengths. Filter length is a trade-off between
    precision and speed, as shorter filters are faster but less precise, and
  longer filters are more precise but computationally more expensive. The red
  square in subplot (b) is the starting point of Figure~\ref{fig:Figure-3}.}%
%

Figure~\ref{fig:Figure-3} shows four consecutive inversion runs for $N=201$,
where each run is a focus on a subsection of the previous run, indicated by the
red square. From the low resolution overview runs (a) and (b) it looks like
this is a standard, straight-forward minimization problem. However, from the
more detailed results in (c) and (d) it becomes obvious that it is a
minimization problem that has to be solved stochastically, as there are
solutions with 2 orders of magnitude difference apparently randomly next to
each other.
%
\plot*{Figure-3}{width=\textwidth}{%
  Solution spaces of four consecutive inversion runs for a 201\,pt filter. Each
  consecutive run zooms into a portion of the previous solution space,
  indicated by the red square. The more in detail we obtain the solution space,
  the more random appears to be the distribution.}%
%

Figure~\ref{fig:Figure-4} shows in (a) the filter values for $J_0$ and $J_1$ of
the best 201 \old{point}\new{pt} filter obtained this way, and in (b) its
right-hand-side solution. The black dots indicate negative values, from which
it can \new{be} seen that adjacent values are often alternating between
positive and negative contributions.
%
\plot*{Figure-4}{width=.9\textwidth}{(a) Filter values of the best obtained
  201\,pt filter with the corresponding check of quality in (b). Black points
  indicate negative values, which shows that adjacent values have often
  opposite signs.}
%

Designing filters is to a large extent trial and error. All input variables
influence the outcome, and often an even better filter is found by sheer luck
of choosing the right starting parameters, as can be seen in
Figure~\ref{fig:Figure-4} (h). However, once you are in this zone every filter
is a very good filter. Each of the input parameters has an effect on the
outcome. It also depends heavily on the transform pairs $f_\mr{I;C}$; functions
that decay rapidly are generally better, as noted by earlier authors
\citep[e.g.][]{USGS.75.Anderson}. It also depends on the filter length, and
obviously heavily on the spacing and shift values. Another important point is
how you define the right-hand evaluation points of the inversion
($r_\mr{def}$). Evaluating corresponding transform pairs separately or jointly
also leads to different filter coefficients ($J_0$, $J_1$, or $J_0$ \& $J_1$,
or equally sine, cosine, or sine \& cosine); also if the real or the imaginary
part is used when complex transform pairs or \texttt{empymod} is used; and if
it is inverted for the minimum amplitude or for the maximum $r$.

\subsection{CSEM}
In this section we compare the 201\,pt filter derived in the previous section
to marine CSEM models \new{as} used in \cite{GP.07.Kong}, \new{in}
\cite{GEO.12.Key}, \old{as well as}\new{and} for a land EM case.

Figure~\ref{fig:Figure-5-bw} compares the derived 201\,pt filter with the two
half-spaces case used by \cite{GP.07.Kong} in his Figure~5. The model consists
of a $0.3125\,\Omega\,$m sea layer of infinite thickness above a subsurface
$1\,\Omega\,$m half-space. The signal of a $1\,$Hz frequency $x$-directed
electric source 50\,m above the interface is measured at an $x$-directed
electric receiver at the interface.
%
\plot*{Figure-5-bw}{width=\textwidth}{(a) Results of different filters and QWE
  for the model of Figure~5 from \cite{GP.07.Kong} with the relative errors
  shown in (b), using the QWE result. The relative error is meaningless from
  about 15.5\,km onwards, as QWE failed as well for these very low amplitudes.}
%
In (a) it can be seen that the new 201\,pt filter is able to recover smaller
amplitudes than the 241\,pt filter from \cite{GP.07.Kong}, the 201\,pt filter
from \cite{GEO.12.Key}, and the 801\,pt filter from \cite{TMS.82.Anderson}
(Wer201, Kong241, Key201, and And801, respectively). It behaves equally well as
the quadrature with extrapolation (QWE), for which we used a 51\,pt quadrature
with relative and absolute tolerance of $10^{-12}$ and $10^{-30}$,
respectively. The relative error is shown in (b), where the QWE result was
taken as \emph{truth}. For offsets greater than roughly 15.5\,km the relative
error becomes meaningless, as the QWE fails itself; this part is greyed out in
the figure.
